---
title: "Bar Charts"
Date: 2021-01-25
output:
  md_document:
    variant: gfm
---

## About

In this notebook, we:

- load the `tidyverse` using `library()`
- load a `csv` file into a `data.frame` using `read.csv()`
- preview a `data.frame` using `head()`
- display a simple bar plot using `ggplot` and `geom_bar`

```{R}
library(tidyverse)
```

## Working with raw data

Load the data with `read.csv`

```{r}
x <- read.csv("../data/tattoos.csv"); head(x)
```

We can see that the data is currently in its raw form.
That is, the scores have not already been aggregated.
When this is the case, we would `geom_bar()` to plot the `Score` column.

```{R}
ggplot(data = x) + 
  geom_bar(
    mapping = aes(x = Score),  # column to plot
    color = "darkblue",  # border colour
    fill = "lightblue",  # fill colour
    width = 0.5) +  # width of each bar
  labs(
    title = "Frequency of Score",  # title of the plot
    x = "Score",  # x-axis label
    y = "Frequency")  # y-axis label
```

## Working with summarised data

Load the data with `read.csv`

```{r}
y <- read.csv("../data/workforce.csv"); head(y)
```

The data could already be summarised, as it is in this case.
Here we would instead use `geom_col` to display the actual data values to plot `Occupation.type` against `Total`.

Note that the `theme()` function is used to adjust how the categorical labels appear on the x-axis.

```{r}
ggplot(data = y) +
  geom_col(
    mapping = aes(x = Occupation.type, y = Total),
    color = "darkblue",
    fill = "lightblue") +
  labs(
    title = "Total UK Workforce by Occupation Type",  # title of the plot
    x = "",  # Suppress the x-axis label
    y = "Frequency (millions)") +  # y-axis label
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1))
```

Note that `R` sorts categorical data alphabetically.
We can change the order of the categories by using the `reorder()` function.

```{r}
ggplot(data = y) +
  geom_col(
    mapping = aes(x = reorder(Occupation.type, -Total), y = Total),
    color = "darkblue",
    fill = "lightblue") +
  labs(title = "Total UK Workforce by Occupation Type",  # title of the plot
    x = "",  # Suppress the x-axis label
    y = "Frequency (millions)") +  # y-axis label
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1))
```

## References

- Activities 4, 5, 6 **(M248 Computer Book A)**
- [Data visualisation](https://r4ds.had.co.nz/data-visualisation.html#statistical-transformations) **(R for Data Science)**
- [Bar charts](https://ggplot2.tidyverse.org/reference/geom_bar.html) **(ggplot2)**
- [Modify axis, legend, and plot labels](https://ggplot2.tidyverse.org/reference/labs.html) **(ggplot2)**
- [Reordering a ggplot bar chart axis](https://www.rpubs.com/dvdunne/reorder_ggplot_barchart_axis) **(RPubs)**
