---
title: "Boxplots"
Date: 2021-01-25
output:
  md_document:
    variant: gfm
---

## About

```{r echo=FALSE}
#import the tidyverse
library(tidyverse)
```

## Boxplots

Load the data using `read.csv()`.

```{r}
# import the data
membership <- read.csv(file = "../data/membership.csv");
head(membership)
```

Plot a boxplot by using `geom_boxplot()` function.
Whether the boxplot is horizontal or vertical depends on what parameter we set in the mapping parameter.

### Horizontal boxplot

```{r}
ggplot(data = membership) +
  geom_boxplot(
    mapping = aes(x = Percentage)) +
  labs(
    title = "Boxplot of Membership"
  )
```

### Vertical boxplot

Plot a default vertical boxplot by using `geom_boxplot()`, setting `aes(y)`.

```{r}
ggplot(data = membership) +
  geom_boxplot(
    mapping = aes(y = Percentage)) +
  labs(
    title = "Boxplot of Membership")
```

## References

- Activities 8, 9 **(M248 Computer Book A)**
- [A box and whiskers plot (in the style of Tukey)](https://ggplot2.tidyverse.org/reference/geom_boxplot.html?q=boxp) **(ggplot2 documentation)**
