---
title: "Frequency histograms"
date: 2021-01-26
output:
  md_document:
    variant: gfm
---

## About

```{r include=FALSE}
library(tidyverse)
```

## Frequency histograms

Load the data with `read.csv`

```{r}
# import the data
membership <- read.csv(file = "../data/membership.csv"); head(membership)
```

Plot a frequency histogram using `geom_histogram`.

```{r}
ggplot(data = membership) +
  geom_histogram(
    mapping = aes(x = Percentage),
    color = "darkblue",
    fill = "lightblue",
    boundary = 0) +  # Bin position specifiers
  scale_x_continuous(
    breaks = seq(16, 30, 2)) +
  labs(
    title = "Frequency Histogram of Membership",
    y = "Frequency")
```

We can alter the number of bins by using the `binwidth=` parameter.

```{r}
ggplot(data = membership) +
  geom_histogram(
    mapping = aes(x = Percentage),
    color = "darkblue",
    fill = "lightblue",
    boundary = 0,  # Bin position specifiers
    binwidth = 1) +  # set the bin width
  scale_x_continuous(
    breaks = seq(16, 30, 2)) +
  labs(
    title = "Frequency Histogram of Membership",
    y = "Frequency")
```

## Relative frequency histograms

Plot a default histogram using `ggplot + geom_histogram`, specifying the parameter `y=..density..`.

```{r}
ggplot(data = membership) +
  geom_histogram(mapping = aes(x = Percentage, y=..density..))
```

## References

- Activities 8, 9 **(M248 Computer Book A)**
- [Histograms and frequency polygons](https://ggplot2.tidyverse.org/reference/geom_histogram.html) **(ggplot2 documentation)**
- [Basic visualization](https://www.sfu.ca/~mjbrydon/tutorials/BAinR/visualize.html#using-ggplot) **(Basic Analytics in R)**
